A user-extensible and adaptable parser architecture

### 2008-05-27 papers/sgai-2008/paper.tex
Pruned some of the content.

Decoupling the parsing rules from the associated actions and framework allows
new rules to be written and tested without requiring modifications to the
parser source code (significantly lowering the barrier to entry for new or
casual users who need to parse new log lines), and greatly simplifies
framework, actions and rules. Decoupling the actions from the
framework simplifies both framework and actions: the framework provides
services to the actions, and doesn't need to deal with the complications which
arise, or the task of reconstructing a mail's journey through Postfix; actions
benefit from having services provided by the framework, freeing them to
concentrate on the task of accurately reconstructing each mail's journey
through Postfix and dealing with the complications arising (discussed in
\sectionref{complications}).

@@

The function tracks the number of times each rule matches; this data is
used during subsequent parsing to optimise the order the rules are tried
in.  Rules specify the Postfix component which produces the log lines they
parse; this avoids needlessly trying rules which won't match the log line,
or worse, might match unintentionally.

@@

Actions contain the code required to deal with log lines from a particular
category, e.g.\ rejections.  The actions are parser-specific: each parser
will need to write the actions it requires from scratch, unless it's
extending an existing parser.  In the Postfix parser developed for this
project there are \numberOFactions{} actions and \numberOFrules{} rules.
The distribution of rules to actions is very skewed, as can be seen in
Fig.~\refwithpage{Distribution of rules per action}.
\showgraph{build/plot-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} The distribution is skewed for
the same reason the architecture is so successful in enabling users to
parse new inputs: most new inputs can be parsed by a combination of a new
rule and an existing action.  Unsurprisingly the action with the highest
number of associated rules is \texttt{REJECTION}, the action which handles
rejections; it's followed by \texttt{INFO}, the action responsible for
handling informative inputs: these are log lines that supplement the data
gathered from other log lines, but that don't themselves represent
interesting events.  The third most common action is, perhaps surprisingly,
\texttt{IGNORE}: there are many log lines which are not of interest to the
parser, e.g.\ warning messages about \DNS{} mismatches, or log lines
lacking a way to identify which data structure they should be associated
with.\footnote{Postfix allocates a queueid (explained in the
glossary,~\sectionref{Glossary}) to each mail; most log lines contain the
queueid of the associated mail, but some log lines, particularly with older
versions of Postfix, lack a queueid, preventing their association with a
particular data structure.}  The parser warns about every unparsed input,
to alert the user that they need to alter or extend their rules, hence the
\texttt{IGNORE} action: the input has been parsed, but it is of no benefit
so it is ignored.  The remaining actions have only one or two associated
rules: in some cases this is because there will only ever be one type of
input for that action, e.g.\ all log lines showing that a remote client has
connected are matched by a single rule and handled by the \texttt{CONNECT}
action; in other cases the action is required to address a deficiency in
the log files, or a complication which arises during
parsing.\footnote{\label{complications footnote}The complications which
arose during development of the parser will be documented in detail in the
author's thesis.}

@@

During development of the parser it became apparent that in addition to the
obvious variety in log lines, there were a large number of complications to
be overcome.\footref{complications footnote}  Some were the result of
deficiencies in Postfix's logging, which were improved or resolved by later
versions of Postfix; others were due to the vagaries of process scheduling,
client behaviour, and administrative actions.  All were successfully
accommodated in the parser: adding new actions was enough for several of
the complications; others required modifications to existing actions to
work around the difficulties; the remainder were resolved by adapting
actions to pass information to subsequent actions (via the framework).

@@

    \item [regex] The \regex{} to match the log line against.  The \regex{}
        will first have several keywords expanded: this simplifies reading
        and writing rules; avoids needless repetition of complex \regex{}
        components; allows the components to be corrected and/or improved
        in one location; and makes each \regex{} largely self-documenting.
        For efficiency the keywords are expanded and every rule's \regex{}
        is compiled before attempting to parse the log file --- otherwise
        each \regex{} would be recompiled each time it was used, resulting
        in a large, data dependent slowdown.  Rule efficiency concerns are
        discussed in \sectionref{rule efficiency}.

    \item [result\_cols] Specifies how the fields in the
        log line will be extracted.  The format is:
        \newline \tab{} \texttt{smtp\_code = 1; recipient = 2, sender = 4;}
        \newline i.e.\ semi-colon or comma separated assignment statements,
        with the variable name on the left and the matching capture from
        the \regex{} on the right hand side.

    \item [result\_data] Sometimes rules need to supply a piece of data
        which isn't present in the log line: e.g.\ setting
        \texttt{smtp\_code} when mail is accepted.  The format and allowed
        variables are the same as for \texttt{result\_cols}, except that
        arbitrary data\footnote{Commas and semi-colons cannot be escaped
        and thus cannot be used.  This is intended for use with small
        amounts of data rather than large amounts in any one rule, so
        dealing with escape sequences was deemed unnecessary.} is permitted
        on the right hand side of the assignment.

@@

    \item [hits] This counter starts at zero and is incremented each time
        the rule successfully matches.  At the start of each run the parser
        sorts the rules in descending order of hits, and at the end of the
        run updates every rule's hits.  Assuming that the distribution of
        log lines is reasonably consistent between log files, rules
        matching more commonly occurring log lines will be tried before
        rules matching less commonly occurring log lines, lowering the
        program's execution time.  Rule ordering for efficiency is
        discussed in \sectionref{rule ordering for efficiency}.

    \item [priority] This is the user-configurable companion to hits: rules
        will be tried in order of priority, overriding hits.  This allows
        more specific rules to take precedence over more general rules
        (described in \sectionref{overlapping rules}).

@@

@@

Scalability, ordering, caching regexes.  Interesting to run with all rules
versus running with the minimum number necessary to parse the test log
files.  Then try that again with different orderings?  Coverage.

@@

MINIMUM VERSUS MAXIMUM NUMBER OF RULES\@.

@@

@@

The dips at log files 22 and 62--68 correspond to peaks in log file size in
graph~\refwithpage{execution time vs file size vs number of lines graph},
and peaks in graph~\refwithpage{execution time vs file size vs number lines
factor} (where a peak means that more lines are processed per second, i.e.\
performance is better).  The explanation for this took some time to arrive
at, but it turns out to be reasonably simple.  The large log files in
question were caused by a mail forwarding loop, where the distribution of
log lines is quite different to normal, resulting in different performance
characteristics.  Fewer

@@

>>>>

Decoupling the parsing rules from the associated actions and framework
allows new rules to be written and tested without requiring modifications
to the parser source code (significantly lowering the barrier to entry for
new or casual users who need to parse new log lines), and greatly
simplifies the framework, actions and rules. Decoupling the actions from
the framework simplifies both framework and actions: the framework provides
services to the actions, and doesn't need to deal with the complications
which arise, or the task of reconstructing a mail's journey through
Postfix; actions benefit from having services provided by the framework,
freeing them to concentrate on the task of accurately reconstructing each
mail's journey through Postfix and dealing with the complications arising
(discussed in \sectionref{complications}).

@@

@@

Actions contain the code required to deal with log lines from a particular
category, e.g.\ rejections.  The actions are parser-specific: each parser
will need to write the actions it requires from scratch, unless it's
extending an existing parser.  In the Postfix parser developed for this
project there are \numberOFactions{} actions and \numberOFrules{} rules.
The distribution of rules to actions is very skewed, as can be seen in
Fig.~\refwithpage{Distribution of rules per action}.
\showgraph{build/plot-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} The distribution is skewed for
the same reason the architecture is so successful in enabling users to
parse new inputs: most new inputs can be parsed by a combination of a new
rule and an existing action.  Unsurprisingly the action with the highest
number of associated rules is \texttt{REJECTION}, the action which handles
rejections; it's followed by \texttt{INFO}, the action responsible for
handling informative inputs: these are log lines that supplement the data
gathered from other log lines, but that don't themselves represent
interesting events.  The third most common action is, perhaps surprisingly,
\texttt{IGNORE}: there are many log lines which are not of interest to the
parser, e.g.\ warning messages about \DNS{} mismatches, or log lines
lacking a way to identify which data structure they should be associated
with.\footnote{Postfix allocates a queueid (explained in the
glossary,~\sectionref{Glossary}) to each mail; most log lines contain the
queueid of the associated mail, but some log lines, particularly with older
versions of Postfix, lack a queueid, preventing their association with a
particular data structure.}  The parser warns about every unparsed input,
to alert the user that they need to alter or extend their rules, hence the
\texttt{IGNORE} action: the input has been parsed, but it is of no benefit,
so it is ignored.  The remaining actions have only one or two associated
rules: in some cases this is because there will only ever be one type of
input for that action, e.g.\ all log lines showing that a remote client has
connected are matched by a single rule and handled by the \texttt{CONNECT}
action; in other cases the action is required to address a deficiency in
the log files, or a complication which arises during
parsing.\footnote{\label{complications footnote}The complications which
arose during development of the parser will be documented in detail in the
author's thesis.}

@@

During development of the parser it became apparent that in addition to the
obvious variety in log lines, there were a large number of complications to
be overcome.\footref{complications footnote}  Some were the result of
deficiencies in Postfix's logging, which were improved or resolved by later
versions of Postfix; others were due to the vagaries of process scheduling,
client behaviour, and administrative actions.  All were successfully
accommodated in the parser: adding new actions was enough for several of
the complications; others required modifications to existing actions to
work around the difficulties; the remainder were resolved by adapting
actions to pass information to subsequent actions (via the framework),
which then check for the extra information and act appropriately.

@@

    \item [regex] The \regex{} to match the log line against.  For
        efficiency every rule's \regex{} is compiled before attempting to
        parse the log file --- otherwise each \regex{} would be recompiled
        each time it was used, resulting in a large, data dependent
        slowdown.\footnote{This is discussed in more detail in the author's
        thesis.}

    \item [result\_cols] Specifies how the fields in the
        log line will be extracted.  The format is:
        \tab{} \texttt{smtp\_code = 1; recipient = 2, sender = 4;}
        \newline i.e.\ semi-colon or comma separated assignment statements,
        with the variable name on the left and the matching capture from
        the \regex{} on the right hand side.

    \item [result\_data] Sometimes rules need to supply a piece of data
        which isn't present in the log line: e.g.\ setting
        \texttt{smtp\_code} when mail is accepted.  The format is the same
        as for \texttt{result\_cols}, except that arbitrary
        data\footnote{Commas and semi-colons cannot be escaped and thus
        cannot be used.  This is intended for use with small amounts of
        data rather than large amounts in any one rule, so dealing with
        escape sequences was deemed unnecessary.} is permitted on the right
        hand side of the assignment.

@@

    \item [hits] This counter is used to optimise the order in which rules
        are tried against the input lines.  Assuming that the distribution
        of log lines is reasonably consistent between log files, rules
        matching more commonly occurring log lines will be tried before
        rules matching less commonly occurring log lines, lowering the
        program's execution time.

    \item [priority] This is the user-configurable companion to hits:
        priority overrides hits when sorting rules.  This allows more
        specific rules to take precedence over more general rules
        (described in \sectionref{overlapping rules}).

@@

XXX MAYBE REMOVE THE SECTION ON DETECTING OVERLAPPING RULES\@?

@@

@@

XXX MINIMUM VERSUS MAXIMUM NUMBER OF RULES\@.

@@

XXX MINIMUM VERSUS MAXIMUM NUMBER OF RULES\@.

@@

The dips at log files 22 and 62--68 correspond to peaks in log file size in
fig.~\refwithpage{execution time vs file size vs number of lines graph},
and peaks in fig.~\refwithpage{execution time vs file size vs number lines
factor} (where a peak means that more lines are processed per second, i.e.\
performance is better).  The explanation for this took some time to arrive
at, but it turns out to be reasonably simple.  The large log files in
question were caused by a mail forwarding loop, where the distribution of
log lines is quite different to normal, resulting in different performance
characteristics.  Fewer

@@

\label{coverage}
